---
title: "lobster_exploration"
author: "guy_fawkes_mask"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
rm(list = rm())
```


```{r}
library(tidyverse)
library(here)
library(janitor)
library(dplyr)
```

# pt 2 or something

## load in csv + -99999 = NA
```{r}
lobster_abundance <- read_csv(here("data", "Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", ""))
```


## feed abundance to cleanup
```{r}
lobster_cleanup <- lobster_abundance
```


## convert column names to lower snake case + uncount

## uncount did row * count value, so now there are "count" number of duplicates in the table
```{r}
lobster_cleanup <- lobster_cleanup %>% clean_names() %>% uncount(count)
```

## lets fucking go

```{r}
count_n_length_smmry <- lobster_cleanup %>% group_by(site, year) %>% summarize(count_by_year = n(), mean_carp = mean(size_mm, na.rm = TRUE))
```

## ggplot + faceting
```{r}
ggplot(data = count_n_length_smmry, aes(x = year, y = count_by_year)) + geom_point(size = 3) + facet_wrap(~ site)
```

## subset to just 2020
```{r}
lobster_2020 <- lobster_cleanup %>% filter(year == "2020")
```

## get counts for lobsters observed at each site in 2020

### add the filled legal column
#### i know i was supposed to use case when but i didnt read that far sorry
#### anyway this still works
```{r}
lobster_2020 <- lobster_2020 %>% mutate(legal = if_else(size_mm >= 79.76, "yes", if_else(size_mm < 79.76, "no", "none")))
```

### subset of legal counts by site + add yes + no somehow
```{r}
lobster_2020_legal <- lobster_2020 %>% group_by(site) %>% count(legal)

lob_2020_pivot_wd <- lobster_2020_legal %>% pivot_wider(names_from = legal, values_from = n)

lobster_2020_legal <- lob_2020_pivot_wd %>% mutate(total = yes + no)
```



### make the stacked column graph
### i computed totals then i just divided them by the yeses
```{r}
ggplot(data = lobster_2020_legal, aes(x = site, y = yes/total)) + geom_col()

```

# pt 3 - practice

## filter()

### ex a
```{r}
ex_a <- lobster_cleanup %>% filter(site == c("IVEE", "CARP", "NAPL"))
```

### ex b
```{r}
ex_b <- lobster_cleanup %>% filter(month == 8)
```

### ex c
```{r}
ex_c <- lobster_cleanup %>% filter(site == "AQUE" | size_mm > 70)

```

### ex d
```{r}
ex_d <- lobster_cleanup %>% filter(site != "NAPL")
```

## group_by() %>% summarize()

### ex_e
```{r}
ex_e <- lobster_cleanup %>% group_by(site) %>% summarize(cara_mean = mean(size_mm, na.rm = TRUE), cara_sd = sd(size_mm, na.rm = TRUE))

```

### ex_f
```{r}
ex_f <- lobster_cleanup %>% group_by(site, month) %>% summarize(max_cara = max(size_mm, na.rm = TRUE))
```

## mutate()

### ex_g
```{r}
ex_g <- lobster_cleanup %>% mutate(length_cm = size_mm * 0.1)

```

### ex_h
```{r}
ex_h <- lobster_cleanup %>% mutate(site = tolower(site))

```

### ex_i
```{r}
ex_i <- lobster_cleanup %>% mutate(area = as.character(area))

```

## case_when()

### ex_j
```{r}
ex_j <- lobster_cleanup %>% 
  mutate(size_bin = case_when(
    size_mm <= 70 ~ "small",
    size_mm > 70 ~ "large"
  )
)

```

### ex_k
```{r}
ex_k <- lobster_cleanup %>% 
  mutate(designation = case_when(
    site == "IVEE" | site == "NAPL" ~ "MPA",
    .default = "not MPA"
  )
)

```
